\documentclass{article}
\input{503macros.tex}

\usepackage{hyperxmp}
\usepackage{hyperref}
%%% FitH fit horizontal, Fit fit whole page, UseNone no bookmarks panel
\hypersetup{pdfpagemode=UseNone,pdfstartview=FitH,pdfpagelayout=OneColumn} 


\begin{document}
\begin{center}
  {\LARGE CSC 503 Homework Assignment 4}\\[1pc]
  Out: September 19, 2018 \\
  Due: September 26, 2018 \\
  \unityid{zzha}
\end{center}


% Unless directed otherwise, follow the convention of the text and
% assume that $a,b,c,d,e$ are constant symbols, $f,g,h$ are function
% symbols, and $w,u,v,x,y,z$ are variable symbols.


\noindent
 Let $a$ and $b$ be constant symbols,
  $f$ a function symbol with three arguments,
  $g$ a function symbol with one argument,
  $h$ a function symbol with two arguments,
  $P$ a predicate symbol with three arguments, and
  $Q$ a predicate symbol with two arguments, and
  $x$, $y$, and $z$ variable symbols.

Indicate, for each of the following strings, which strings are
formulas in predicate logic, and state a reason for failure for
strings which are not.  No credit will be given for strings correctly
identified as not being formulas without correct identification of a
reason why the string is not a formula.

\begin{enumerate}
\item[1.] \textbf{[5 points]} 
    $\exists y\ P(x,b,y) \land P(h(b,x),h(y,b),g(y)(a))$

  \begin{answer}
    Not a formula, $g(y)(a)$ is not a valid term.
  \end{answer}
  \end{enumerate}

\begin{enumerate}
\item[2.] \textbf{[5 points]}
  $\exists b\ \forall x\ Q(f(b,y,x),g(h(a,x)))$

  \begin{answer}
    Not a formula, $b$ is not a variable where after the existential quatifier.
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[3.] \textbf{[5 points]}
    $\forall z\ \exists x\ P(g(h(b,x),z),f(a,y))$
    
  \begin{answer}
    Not a formula, $P$ needs 3 arguments.
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[4.] \textbf{[5 points]}
  $\forall y\ g(Q(z,y)) \implies P(h(f(y,z,z),z))$
  
  \begin{answer}
    Not a formula, implies only apply to valid formula, $g(Q(z,y))$ is not a valid formula.
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[5.] \textbf{[5 points]}
  $P(a,g(g(f(a,\neg b, a))),a) \implies P(a,f(a,\neg b, a),a)$
    
  \begin{answer}
    Yes, it is a valid formula.
  \end{answer}
\end{enumerate}

\noindent
Now let $R$ be a predicate symbol with arity 2, $f$ be a function of
arity 2, and $\phi$ be the formula
\begin{displaymath}
  \forall x\
  [(R(x,z) \land  \exists z\ \neg R(z,f(x,y)))
  \implies \forall y\ R(y,z)]
\end{displaymath}
  
\begin{enumerate}
\item[6.] \textbf{[5 points]} Indicate, for each occurrence of each
  variable in $\phi$, whether that occurrence is free or bound.

  \begin{answer}
    \begin{displaymath}
      \forall x\
      [(R(x_{bound},z_{free}) \land  \exists z\ \neg R(z_{bound},f(x_{bound},y_{free})))
      \implies \forall y\ R(y_{bound},z_{free})]
    \end{displaymath}
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[7.] \textbf{[5 points]} List all variables which occur both free and
  bound in $\phi$.
  \begin{answer}
    \begin{enumerate}
      \item $z$
      \item $y$
    \end{enumerate}
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[8.] \textbf{[5 points]} Compute $\phi[t/x]$ for
  $t = h(f(g(y),x),a,x)$.  Is $t$ free for $x$ in $\phi$?
  \begin{answer}
    Yes, $t$ is free for $x$ in $\phi$. Because all the occurrences of $x$ are bounded, 
    thus no substitution is performed, and therefor no free $x$ leaf in $\phi$ occurs 
    in the scope of $\forall y$ or $\exists y$ for any variable $y$ occurring in $t$.
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[9.] \textbf{[5 points]} Compute $\phi[t/y]$ for
  $t = h(f(g(y),x),a,x)$.  Is $t$ free for $y$ in $\phi$?
  \begin{answer}
    No, $t$ is not free for $y$ in $\phi$. Because the free $y$ is bound in the scope 
    of $\forall x$, where variable $x$ occurring in $t$ .
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[10.] \textbf{[5 points]} Compute $\phi[t/z]$ for
  $t = h(f(g(y),x),a,x)$.  Is $t$ free for $z$ in $\phi$?
  \begin{answer}
    No, $t$ is not free for $z$ in $\phi$. Because the free $z$ is bound in the scope 
    of $\forall x$ and $\forall y$, where variable $x$ and $y$ occurring in $t$ .
  \end{answer}
\end{enumerate}

\noindent
Now consider a language in which the only nonlogical symbols are a
predicate symbol $R$ of two arguments and a function symbol $f$ of one
argument.  Let $\phi_1$ and $\phi_2$ be the sentences
\begin{eqnarray*}
  \phi_1 
  & =
  & \forall x\ \forall y\ \forall z\
    R(x,y) \land R(y,z) \implies R(x,z) 
  \\
  \phi_2
  & = 
  & \forall x\ \forall y\ \forall z\
    R(x,y) \implies R(f(x,z),f(y,z))
\end{eqnarray*}
Recall that an interpretation of the language identifies a domain of
interpretation $M$ (a set), and maps each of the predicate and
function symbols to a set of tuples of elements of $M$ of the
appropriate size.

\begin{enumerate}
\item[11.] \textbf{[15 points]:} Give a formal interpretation $I$ that
  makes $\phi_1$ true and $\phi_2$ false.

  \begin{answer}
    \begin{enumerate}
      \item Domain $M = \{a, b\}$
      \item Function $f(a,a) = b$; $f(a,b) = b$; $f(b,a) = b$; $f(b,b) = b$;
      \item Interpretation $R^I = \{(a, a)\} $
    \end{enumerate}
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[12.] \textbf{[5 points]:} Briefly explain why $I$ makes $\phi_1$
  true.
  \begin{answer}
    In order to make for all $x, y, z$, the $\phi_1$ being true:
    \begin{enumerate}
      \item Case 1: if $x = y = z = a$, then $R(x,y), R(y,z), R(x,z)$ all are true.
      \item Other than Case 1: any $x, y, z$ combination will result a case either 
      $R(x,y)$ or $R(y,z)$ being false, which makes the whole being true.
    \end{enumerate}
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[13.] \textbf{[5 points]:} Briefly explain why $I$ makes $\phi_2$
  false.
  \begin{answer}
    In order to make for all $x, y, z$, the $\phi_2$ being false, just need to find 
    there exist one false case:
    
    if $x = a, y = a$, then $R(x,y)$ is true, and $R(f(x,z),f(y,z))$ would be $R(b, b)$ 
    which is false, therefore makes the $\phi_2$ being false.
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[14.] \textbf{[15 points]:} Give the formal definition of an
  interpretation $J$ that makes $\phi_1$ false and $\phi_2$ true.
  
  \begin{answer}
    \begin{enumerate}
      \item Domain $M = \{a, b\}$
      \item Function $f(a,a) = a$; $f(a,b) = a$; $f(b,a) = b$; $f(b,b) = b$;
      \item Interpretation $R^I = \{(a, b), (b, a)\} $
    \end{enumerate}
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[15.] \textbf{[5 points]:} Briefly explain why $J$ makes $\phi_1$
  false.
  \begin{answer}
    In order to make for all $x, y, z$, the $\phi_1$ being false, just need to find 
    there exist one false case:
    
    if $x = a, y = b, z = a$, then $R(x,y)$ and $R(y,z)$ is true, and $R(x,z)$ is 
    false, therefore makes the $\phi_1$ being false.
  \end{answer}
\end{enumerate}

\begin{enumerate}
\item[16.] \textbf{[5 points]:} Briefly explain why $J$ makes $\phi_2$
  true.
  \begin{answer}
    In order to make for all $x, y, z$, the $\phi_2$ being true:
    \begin{enumerate}
      \item Case 1: if $x = y $, then $R(x,y)$ is either $R(a,a)$ or $R(b, b)$, 
      which is false, makes the whole being true.
      \item Other 2: $x \neq y $, then $R(f(x,z),f(y,z))$ would be either $R(a,b)$ 
      or $R(b, a)$ which is true in either case, which makes the whole being true.
    \end{enumerate}
  \end{answer}
\end{enumerate}

\end{document}
