\documentclass{article}
\input{503macros.tex}

\begin{document}
\begin{center}
  {\LARGE CSC 503 Homework Assignment 3}\\[1pc]
  Out: September 10, 2018 \\
  Due: September 17, 2018 \\
  \unityid{MISSING}
\end{center}

\begin{enumerate}

\item \textbf{[30 points]} Using the method described in lecture, construct a
  formula $\phi$ in \textbf{DNF} to match the following truth table.
  Show and explain any intermediate steps.
  \begin{center}
    \begin{tabular}{ccc|c}
      $\alpha$ & $\beta$ & $\gamma$ & $\phi$ \\ \hline
      T & T & T & T\\
      T & T & F & T\\
      T & F & T & F\\
      T & F & F & T\\
      F & T & T & T\\
      F & T & F & F\\
      F & F & T & F\\
      F & F & F & T
    \end{tabular}
  \end{center}

\item \textbf{[30 points]} Using the method described in lecture,
  construct a formula $\psi$ in \textbf{CNF} to match the following
  truth table.  Show and explain any intermediate steps.
  \begin{center}
    \begin{tabular}{ccc|c}
      $\alpha$ & $\beta$ & $\gamma$ & $\psi$ \\ \hline
      T & T & T & T\\
      T & T & F & F\\
      T & F & T & F\\
      T & F & F & T\\
      F & T & T & T\\
      F & T & F & T\\
      F & F & T & F\\
      F & F & F & T
    \end{tabular}
  \end{center}

  \vfill
  \begin{center}
    (continued overleaf)
  \end{center}
  \newpage
  
\item \textbf{[30 points total]} Apply the following version of the
  algorithm HORN from pages 66--67 of the textbook to the following
  Horn formula $\varphi$.
  \begin{enumerate}
  \item Mark each occurrence of $\top$ in each conjunct of $\varphi$
    with a ``1'' to indicate $\top$ is marked on the first pass.
  \item Search the list of conjuncts of $\varphi$ in order until
    either
    \begin{enumerate}
    \item the end of the list is reached, or
    \item a ``markable'' conjunct is found, that is, one with each of
      its antecedent propositions marked and its consequent
      proposition unmarked.
    \end{enumerate}
  \item If the end of the list was reached,
    \begin{enumerate}
    \item If $\bot$ is not marked, return ``satisfiable'' and halt.
    \item If $\bot$ is marked, return ``unsatisfiable'' and halt.
    \end{enumerate}
  \item Otherwise, mark every occurrence of the consequent proposition
    of the first markable conjunct with the pass number (2, 3, \dots)
    in each conjunct of $\varphi$.
  \item Continue with step (b) again.
  \end{enumerate}
  \begin{displaymath}
    \varphi = 
    \begin{array}{|lll}
      1.  & (\top \implies w)                  & \land \\
      2.  & (w \implies q)                     & \land \\
      3.  & (x \land t \implies \bot)          & \land \\
      4.  & (q \land r \implies p)             & \land \\
      5.  & (v \implies s)                     & \land \\
      6.  & (w \implies r)                     & \land \\
      7.  & (r \land s \implies x)             & \land \\
      8.  & (\top \implies v)                  & \land \\
      9.  & (v \land q \implies u)             & \land \\
      10. & (p \land r \land s \implies u)     & \land \\
      11. & (u \implies v)                     &       \\
    \end{array}
  \end{displaymath}
  Your answer should list propositional letters in the order in which
  they are marked and indicate the returned value.

  \item \textbf{[10 points]} Can one determine whether the following
    formula $\psi$ is satisfiable more quickly than by
    applying the HORN algorithm?  Explain your answer.
  \begin{displaymath}
    \psi = 
    \begin{array}{|lll}
        & (\top \implies p)                  & \land \\
        & (\bot \implies q)                  & \land \\
        & (p \land q \implies v)             & \land \\
        & (q \implies r)                     & \land \\
        & (v \implies s)                     & \land \\
        & (\top \implies r)                  & \land \\
        & (\top \implies s)                  & \land \\
        & (v \land t \implies u)             & \land \\
        & (p \land r \land s \implies u)     & 
    \end{array}
  \end{displaymath}

\end{enumerate}
\end{document}
